---
title: "Statistical assignment 4"
author: "Thea Smith"
date: "26th February 2020"
output: github_document
---

```{r setup, include=FALSE}
# Please note these options.
# This tells R Markdown that we want to show code in the output document.
knitr::opts_chunk$set(echo = TRUE)
# Switching off messages in the output document. 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
# Switching on caching to make things faster (don't commit cache files on Github).
knitr::opts_chunk$set(cache = TRUE)
```

In this assignment you will need to reproduce 5 ggplot graphs. I supply graphs as images; you need to write the ggplot2 code to reproduce them and knit and submit a Markdown document with the reproduced graphs (as well as your .Rmd file).

First we will need to open and recode the data. I supply the code for this; you only need to change the file paths.
    ```{r}
library(tidyverse)
Data8 <- read_tsv("/Users/TheaSmith/Desktop/Data Analysis III/assignment-4-tkcs201/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
Data8 <- Data8 %>%
        select(pidp, h_age_dv, h_payn_dv, h_gor_dv)
Stable <- read_tsv("/Users/TheaSmith/Desktop/Data Analysis III/assignment-4-tkcs201/UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")
Stable <- Stable %>%
        select(pidp, sex_dv, ukborn, plbornc)
Data <- Data8 %>% left_join(Stable, "pidp")
rm(Data8, Stable)
Data <- Data %>%
        mutate(sex_dv = ifelse(sex_dv == 1, "male",
                           ifelse(sex_dv == 2, "female", NA))) %>%
        mutate(h_payn_dv = ifelse(h_payn_dv < 0, NA, h_payn_dv)) %>%
        mutate(h_gor_dv = recode(h_gor_dv,
                         `-9` = NA_character_,
                         `1` = "North East",
                         `2` = "North West",
                         `3` = "Yorkshire",
                         `4` = "East Midlands",
                         `5` = "West Midlands",
                         `6` = "East of England",
                         `7` = "London",
                         `8` = "South East",
                         `9` = "South West",
                         `10` = "Wales",
                         `11` = "Scotland",
                         `12` = "Northern Ireland")) %>%
        mutate(placeBorn = case_when(
                ukborn  == -9 ~ NA_character_,
                ukborn < 5 ~ "UK",
                plbornc == 5 ~ "Ireland",
                plbornc == 18 ~ "India",
                plbornc == 19 ~ "Pakistan",
                plbornc == 20 ~ "Bangladesh",
                plbornc == 10 ~ "Poland",
                plbornc == 27 ~ "Jamaica",
                plbornc == 24 ~ "Nigeria",
                TRUE ~ "other")
        )
    ```

Reproduce the following graphs as close as you can. For each graph, write two sentences (not more!) describing its main message.  

1. Univariate distribution (20 points).

    ```{r}
    Data %>%
  ggplot(Data, mapping = aes(x = h_payn_dv)) +
    geom_freqpoly(size = 0.3) + 
    labs(x = "Net Pay Per Month",
         y = "Freuency")
    ```
We see from this chart that the most common net pay per month is about £1700, with about 2600 respondants recieving this. From this point, as the pay increases, the proportion of the population recieving this decreases, with an anamalous increase at about £5500. 

2. Line chart (20 points). The lines show the non-parametric association between age and monthly earnings for men and women.

    ```{r}
AgeEarn <- Data %>%
  group_by(h_age_dv) %>%
  summarise(
    meanIncome = mean(h_payn_dv, na.rm = TRUE)
  )
AgeEarn %>%
  ggplot(aes(x = h_age_dv, y = meanIncome)) +
  geom_line() +
  geom_smooth() +
  xlim(21,80) +
  xlab("Age") +
  ylab("Mean income")

AgeSexEarn <- Data %>%
  filter(!is.na(sex_dv)) %>%
  group_by(h_age_dv, sex_dv) %>%
  summarise(meanIncome = mean(h_payn_dv, na.rm = TRUE))
    
AgeSexEarn %>%
  ggplot(aes(x = h_age_dv, y = meanIncome, colour = sex_dv)) +
  geom_line() +
  geom_smooth() +
  xlim(21,80) +
  ylim(0,2500) +
  xlab("Age") +
  ylab("Mean income")
    ```

This graph shows that though men and women start by earning a very similar abount in their early 20s, men very quickly earn much more than women on average, with a peak of about £2250per month In their 40s while women only peak at about £1500 per month around age 50. 

3. Faceted bar chart (20 points).

    ```{r}
PayBorn <- Data %>%
  filter(!is.na(h_payn_dv)) %>%
  filter(!is.na(placeBorn)) %>%
  group_by(placeBorn, sex_dv) %>%
  summarise(meanIncome = mean(h_payn_dv, na.rm = TRUE))
ggplot(PayBorn, mapping = aes(x = sex_dv, y = meanIncome)) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~ placeBorn) +
  labs(x = "Sex", 
       y = "Mean Net Pay Per Month")
    ```
We can see that in all countries of birth in this data set, the men on average have a higher net pay per month, this could be the result of patriarchal pay gaps. We can also see that the population of some countries are paid more on average, with Ireland having the highest pay and Bangladesh haveing the lowest, possibly as a result of the different opportunities both of education and job positions on offer to people from different countries. 


4. Heat map (20 points).

    ```{r}
 HeatBorn <- Data %>%
  filter(!is.na(h_gor_dv)) %>%
  filter(!is.na(placeBorn)) %>%
  group_by(h_gor_dv, placeBorn) %>%
  summarise(meanAge = mean(h_age_dv, na.rm = TRUE))
ggplot(HeatBorn, mapping = aes(x = h_gor_dv, y = placeBorn, fill = meanAge)) +
  geom_tile() + 
  labs(x = "Government Office Region", 
       y = "Country of birth",
       fill = "Mean age") +
  theme(axis.text.x = element_text(angle = 50))
    ```
From this heat map we can see that generally speaking, regardless of the location, populations from specific countries of birth have similar mean ages, however Poland and Nigeria do fluctuate. this could be as a result of populations from certain countries migrating to the UK at certain times. 

5. Population pyramid (20 points).

range(Data$h_age_dv)

    ```{r}
Data <- Data %>%
  mutate(AgeGroup = case_when(
    between(h_age_dv, 0, 5) ~ '0-5',
    between(h_age_dv, 6, 10) ~ '6-10',
    between(h_age_dv, 11, 15) ~ '11-15',
    between(h_age_dv, 16, 20) ~ '16-20',
    between(h_age_dv, 21, 25) ~ '21-25',
    between(h_age_dv, 26, 30) ~ '26-30',
    between(h_age_dv, 31, 35) ~ '31-35',
    between(h_age_dv, 36, 40) ~ '36-40',
    between(h_age_dv, 41, 45) ~ '41-45',
    between(h_age_dv, 46, 50) ~ '46-50',
    between(h_age_dv, 51, 55) ~ '51-55',
    between(h_age_dv, 56, 60) ~ '56-60',
    between(h_age_dv, 61, 65) ~ '61-65',
    between(h_age_dv, 66, 70) ~ '66-70',
    between(h_age_dv, 71, 75) ~ '71-75',
    between(h_age_dv, 76, 80) ~ '76-80',
    between(h_age_dv, 81, 85) ~ '81-85',
    between(h_age_dv, 86, 90) ~ '86-90',
    between(h_age_dv, 91, 95) ~ '91-95',
    between(h_age_dv, 96, 100) ~ '96-100', 
    between(h_age_dv, 101, 105) ~ '101-105'
  ))


Pop <- Data %>%
        group_by(AgeGroup) %>%
        filter(!is.na(sex_dv)) %>%
        count(sex_dv)

Pop %>%
    ggplot(
    aes(x = AgeGroup, y = ifelse(sex_dv == "female", n, -n), fill = sex_dv)) +
    geom_bar(stat = "identity", width =  1) +
    coord_flip() +
    ylab("Population") +
    xlab("Age")+
    ggtitle("Population Pyramid")
    ```
    
By looking at this population pyramid we can see that overall there are more women than men in this population. We can see that the population decreases towards the 26-30 age before increasing towards the 46-50 age group. The dearease (especially of men) could be down to emmigration of the younger generation and the narrower base indicates an aging population. 